Disease progression of 213 patients hospitalized with Covid-19 in the Czech Republic in March - October 2020: An exploratory analysis

### 2021-06-14 manuscript/manuscript.Rmd
Tiny tweaks before reformatting

At the time the study was conducted, the proposed treatments included antivirals approved for other indications (chloroquine, hydroxychloroquine, lopinavir/ritonavir, remdesivir, favipiravir, umifenovir), azithromycin, corticosteroids, immunoglobulins, tocilizumab and convalescent plasma [@http://zotero.org/users/5567156/items/8XWTUMUI; @http://zotero.org/users/5567156/items/6Q8Z8VJR]. Notably, the anti-malarial and anti-rheumatic drug hydroxychloroquine and the macrolide antibiotic  azithromycin showed promise in early data and were broadly available and thus were frequently used in the early stages of the pandemic. Remdesivir, previously designed and approved for Ebola, SARS and MERS, also reported good initial results. However, during spring and summer 2020 remdesivir was available in Czech Republic only in limited amounts via compassionate use programme. The RECOVERY trial reported positive results of croticosteriod dexamethasone for severe cases in June 2020 [@http://zotero.org/users/5567156/items/W8KK3TX2], but at this point the number of Covid-19 patients hospitalized in Czech Republic was low and dexamethasone thus did not see wider use until later in the pandemic.

@@

A convenience sample of patients from `r n_hospitals` sites was collected. The study sites span the whole spectrum of sizes from large university hospitals in major cities with multiple dedicated Covid-19 wards (Thomayer Hospital in Prague, Motol University Hospital in Prague,  Kralovské Vinohrady University Hospital in Prague, General University Hospital in Prague, University Hospital in Pilsen) through major regional/specialized hospitals  (Na Homolce Hospital in Prague, Military Hospital Olomouc) as well as smaller hospitals caring for just several Covid-19 patients at a time (AGEL Hospital Nový Jičín, Hořovice Hospital, Třebíč Hospital). The sites were chosen based on availability and willingness of the personnel to participate in data collection. None of the study sites was exclusively dedicated to treating Covid-19 patients. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic in March 2020 (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave" (admitted after September 1st). Last patient included in the dataset was admitted on October 12th. See Supplementary Figure 1 for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.

>>>>

At the time the study was conducted, the proposed treatments included antivirals approved for other indications (chloroquine, hydroxychloroquine, lopinavir/ritonavir, remdesivir, favipiravir, umifenovir), azithromycin, corticosteroids, immunoglobulins, tocilizumab and convalescent plasma [@http://zotero.org/users/5567156/items/8XWTUMUI; @http://zotero.org/users/5567156/items/6Q8Z8VJR]. Notably, the anti-malarial and anti-rheumatic drug hydroxychloroquine and the macrolide antibiotic  azithromycin showed promise in early data and were broadly available and thus were frequently used in the early stages of the pandemic. Remdesivir, previously designed and approved for Ebola, SARS and MERS, also reported good initial results. However, during spring and summer 2020 remdesivir was available in Czech Republic only in limited amounts via compassionate use programme. The RECOVERY trial reported positive results of coroticosteriod dexamethasone for severe cases in June 2020 [@http://zotero.org/users/5567156/items/W8KK3TX2], but at this point the number of Covid-19 patients hospitalized in Czech Republic was low and dexamethasone thus did not see wider use until later in the pandemic.

@@

A convenience sample of patients from `r n_hospitals` sites was collected. The study sites span the whole spectrum of sizes from large university hospitals in major cities with multiple dedicated Covid-19 wards (Thomayer University Hospital in Prague, Motol University Hospital in Prague,  Kralovské Vinohrady University Hospital in Prague, General University Hospital in Prague, University Hospital in Pilsen) through major regional/specialized hospitals  (Na Homolce Hospital in Prague, Military Hospital Olomouc) as well as smaller hospitals caring for just several Covid-19 patients at a time (AGEL Hospital Nový Jičín, Hořovice Hospital, Třebíč Hospital). The sites were chosen based on availability and willingness of the personnel to participate in data collection. None of the study sites was exclusively dedicated to treating Covid-19 patients. For each site, the dataset contains all patients hospitalized in the participating wards over the data collection period. The data collection started at the onset of the Covid-19 pandemic in March 2020 (except for one site where some older records were inaccessible), but the end date for collection differs between sites due to time constraints of the participating physicians. Three sites included total of `r n_second_wave` patients that could be considered part of "second wave" (admitted after September 1st). Last patient included in the dataset was admitted on October 12th. See Supplementary Figure 1 for per-site data collection periods. Patients over the age of 18 were included if they had PCR-confirmed infection of SARS-CoV-2 and were not participating in a clinical trial of any Covid-19 pharmacotherapy.
